
Thus far we have used the Coq type \li{nat}, which represents 
a number~$n$ using~$n$ bits. 
The binary system represents~$n$ in $\lfloor \log_{2} n \rfloor + 1$ bits.
In both systems, addition/subtraction of $b$-bit 
numbers is $\Theta(b)$, while multiplication is $\Theta \big(b^2\big)$. 
Arithmetic operations are therefore faster when the inputs 
are encoded in binary\footnote{We put a further explanation of 
Coq's binary type \li{N} in Appendix~\ref{apx:bin_in_coq}.}.

Our codebase has binary versions of hyperoperations, inverse hyperoperations, 
Ackermann, and inverse Ackermann. 
Here we show how to compute inverse Ackermann for binary inputs in 
linear time up to the size of input representation, 
which is equivalent to logarithmic time up to input magnitude.
As before, we present an intuitive sketch here and put 
full proofs in Appendix~\ref{apx:time_analysis_bin}.

\subsection{Countdown and contractions with binary numbers}

Although the theoretical \emph{countdown} is independent of the encoding 
of its inputs, its Coq definition needs to be adjusted to allow for inputs 
in \li{N}. The first step is to translate the arguments of 
\li{countdown\_worker} from \li{nat} to \li{N}. Budget \li{b} must 
remain in \li{nat} so it can serve as Coq's termination argument, 
but all other \li{nat} arguments should be changed 
to \li{N}, and functions on \li{nat} to functions on \li{N}.
\begin{lstlisting}
Fixpoint bin_cdn_wkr (f : N -> N) (a n : N) (b : nat) : N :=
  match b with O => 0 | S b' => 
    if (n <=? a) then 0 else 1 + bin_cdn_wkr f a (f n) b'
  end.
\end{lstlisting}

Determining the budget for \li{bin\_countdown} is tricky. 
A naïve approach is to use the \li{nat} translation of \li{n}, 
\emph{i.e.} \li{N.to_nat n}. This is untenable as it alone 
takes exponential time \emph{viz} the length of $n$'s representation.
We need a linear-time budget calculation for countdowns 
of oft-used functions like $\lambda n.(n-2)$.

The key is to focus on functions that can bring their arguments below a threshold via repeated application in 
\emph{logarithmic} time, thus allowing a log-sized budget for 
\li{bin\_countdown\_worker}. Simply shrinking by~$1$ is no longer good enough;
we need to halve the argument on every application as shown below:
\begin{defn} \label{defn: bin-contraction}
	$f\in \contract$ is \emph{binary strict above} $a\in \mathbb{N}$ if $\forall n > a, f(n) \le \lfloor \frac{n + a}{2} \rfloor$.
\end{defn}
The key advantage of binary strict contractions is that if a contraction $f$ is binary strict above some $a$, then 
$\forall n > a, \forall k.~f(n) \le \left\lfloor \frac{n - a}{2^k} \right\rfloor + a$.
Therefore within $\lfloor \log_2 (n - a) \rfloor + 1$ applications of $f$, the result will become equal to or less than $a$, so we can choose this number as the budget for \li{bin\_countdown\_worker} to successfully reach the countdown value before terminating. 
Note that this budget is simply the length of the binary representation 
of \li{n - a}, which we calculate using our function 
\href{www.example.com}{\li{nat\_size}}.
The Coq definition of countdown on \li{N} is:
\hide{
\begin{lstlisting}
Definition nat_size (n : N) : nat :=
  match n with
  | 0 => 0%nat
  | Npos p => let fix nat_pos_size (x : positive) : nat :=
                  match x with
                  | xH => 1%nat
                  | xI y | xO y => S (nat_pos_size y) 
                  end
              in nat_pos_size p
  end.
\end{lstlisting}
Note that \li{nat_size} outputs $0$ on $0$, and on any positive number $m$ the size of its binary representation, thus equals to $\lfloor \log_{2} m \rfloor + 1$.
}% end hide
\begin{lstlisting}
Definition bin_cdn (f : N -> N) (a n : N) : N := 
  bin_cdn_wkr f a n (nat_size (n - a)).
\end{lstlisting}

\noindent The following is the binary version of Lemma~\ref{lem: cdt-runtime}:
\begin{lem} \label{lem: cdt-runtime-bin}
	$\forall n \in \li{N}$, if $f$ is a binary strict contraction above $a$,
	\begin{equation*}
	\runtime_{\cdt{f}{a}}(n) \le \sum_{i=0}^{\cdt{f}{a}(n) - 1} \hspace{-6pt}
	\runtime_f\big(f^{(i)}(n)\big) \ + \ (\log_2a + 3)\left(\cdt{f}{a}(n) + 1\right) \ + \ 2\log_2n \ + \ \log_2\cdt{f}{a}(n)
	\end{equation*}
\end{lem}

%Substituting $a=1$ into \eqref{eq: cdt-runtime-bin} shows that \Cref{lem: inv-ack-hier-runtime} still holds.
%Similar to \Cref{sect: hardcode-lvl2}, the use of binaries is not immediately effective since the first level. We delve deeper into the hierarchy by \emph{hardcoding the $3^{\text{th}}$ level} and starts from there. Now
%\begin{equation*}
%\forall n, n+2  < n+3 < 2(n+2) \iff \forall n,
%\lfloor \log_2(n+2) \rfloor < \lceil \log_2(n+3) \rceil \le \lfloor \log_2(n+2) \rfloor + 1
%\end{equation*}
%This shift from floor to ceiling enables a direct computation, since $\lceil \log_2(n+3) \rceil$ can now be seen as the size of $(n+2)$'s binary representation.
%\begin{lstlisting}
%Definition alpha3 (n: N) : N := N.size (n+2) - 3.
%\end{lstlisting}
%Let $n\ge 2$. The computation of \li{N.size(n)} takes time equal to itself, so the above definition gives $\runtime(\alpha_3, n) \le 2\log_2n$. Fix an $i\ge 3$ and suppose $\runtime(\alpha_i, n) \le C_i\log_2n$. By \Cref{lem: inv-ack-hier-runtime},
\subsection{Inverse Ackermann in $\Theta\left(\log_2 n\right)$ time}
Our new Coq definition computes countdown 
only for strict binary contractions. Fortunately, starting 
from $n = 2$, the inverse hyperoperations $a\angle{n}b$ when $a\ge 2$ 
and the inverse Ackermann hierarchy $\alpha_n$ are all strict binary contractions.
We can construct these hierarchies by hardcoding their 
first three levels and recursively building higher levels with \li{bin\_countdown}. 
Furthermore, 
analagously to the optimization for \li{nat}, we hardcode an additional level.
\begin{lstlisting}
Fixpoint bin_alpha (m : nat) (x : N) : N :=
  match m with
  | 0%nat => x - 1          | 1%nat => x - 2
  | 2%nat => N.div2 (x - 2) | 3%nat => N.log2 (x + 2) - 2
  | S m'  => bin_cdn (bin_alpha m') 1 (bin_alpha m' x)
  end.
\end{lstlisting}
Note that for all $x$, $\li{N.div2}(x - 2) = \left\lfloor \frac{x - 2}{2} \right\rfloor = \left\lceil \frac{x - 3}{2} \right\rceil$ and $\li{N.log2}(x + 2) - 2 = \left\lfloor \log_2(x+2) \right\rfloor - 2 = \left\lceil \log_2(x+3) \right\rceil - 3$, so the above Coq code is correct. 
 
\begin{thm} \label{thm: inv_ack_runtime_bin}
	$\forall i, \forall n, \runtime_{\alpha_i}(n) \le 2\log_2n + \left(3\cdot 2^i - 3i - 13\right)\log_2\log_2n + 3i$.
\end{thm}

\noindent For any level of the Ackermann hierarchy, this theorem demonstrates
a linear runtime up to the size of the representation of the input.
That is to say, a logarithmic runtime up to its magnitude $n$:
$\runtime_{\alpha_i}(n) = \Theta\big(\log_2n + 2^i\log_2\log_2n \big)$.

Moving on to inverse Ackermann itself, we follow a style nearly identical to that 
in~\S\ref{subsec: inv_ack_hier}. For the worker, we simply translate to 
\li{N}, keeping the budget in \li{nat} as described earlier. 
The Ackermann function has an extra hardcoded level.
\begin{lstlisting}
Fixpoint bin_ia_wkr (f : N -> N) (n k : N) (b : nat) : N :=
  match b with 0%nat  => k | S b' => if n <=? k then k else
    let g := (bin_cdn f 1) in bin_ia_wkr (compose g f) (g n) (N.succ k) b'
  end.

Definition bin_inv_ack (n : N) : N :=
  if (n <=? 1) then 0 else if (n <=? 3) then 1
    else if (n <=? 7) then 2 else
      let f := (bin_alpha 3) in bin_ia_wkr f (f n) 3 (nat_size n).
\end{lstlisting}
% Fixpoint bin_inv_ack_worker (f : N -> N) (n k : N) (b : nat) : N :=
%   match b with
%   | 0%nat  => k
%   | S b' => if n <=? k then k
%             else let g := (bin_countdown f 1) in
%                  bin_inv_ack_worker (compose g f) (g n) (N.succ k) b'
%   end.

% Definition bin_inv_ack (n : N) : N :=
%   if (n <=? 1) then 0
%   else if (n <=? 3) then 1
%        else if (n <=? 7) then 2
%             else let f := (bin_alpha 3) in
%                  bin_inv_ack_worker f (f n) 3 (nat_size n).
Note that, for $n > 7$, $n < \Ack\big(\lfloor \log_2n \rfloor + 1\big)$ $= \Ack\big(\li{nat_size}(n)\big)$, so a budget of $\li{nat_size}(n)$ suffices. 
We prove the correctness of \li{bin_inv_ack} in our codebase.

As in~(\ref{eq: inv-ack-runtime-improved}), summing the time complexity
of each component \li{bin_inv_ack} shows that it is linear up to the 
size of input representation.
\begin{equation*}
\runtime_\alpha(n) \ =
\begin{aligned}
& \sum_{k = 3}^{\alpha(n) - 1} \runtime_{\cdt{(\alpha_k)}{1}\ }(\alpha_k(n)) \ + \ \sum_{k = 3}^{\alpha(n)}\Tleb\left(\alpha_k(n), k\right) \ + \
\sum_{k = 3}^{\alpha(n) - 1}\Tsucc(k)\\
& \ + \runtime_{\alpha_2}(n) \ + \runtime_{\li{nat_size}}(n)
\ + \Tleb(n, 1) \ + \Tleb(n, 3) \ + \Tleb(n, 7)
\end{aligned}
\end{equation*}
In the first summand, $\forall k.~\runtime_{\cdt{(\alpha_k)}{1}\ }(\alpha_k(n)) = \runtime_{\alpha_{k+1}}(n) - \runtime_{\alpha_k}(n)$.
Each $\Tleb$ in the second summand is $\Theta\left(\log_2k\right)$, totalling $\bigO\big(\alpha(n)\log_2\alpha(n)\big)$.
The third summand is $\Theta(\alpha(n))$ by Lemma~\ref{lem: succ-runtime-bin}. The remaining items total $\Theta(\log_2n)$. Therefore,
\begin{equation*}
\begin{aligned}
\runtime_\alpha(n)
& \textstyle = \sum_{k = 2}^{\alpha(n) - 1} \big(\runtime_{\alpha_{k+1}}(n) - \runtime_{\alpha_{k}}(n)\big) + \Theta\left(\log_2 n\right)
= \runtime_{\alpha_{\alpha(n)}}(n) + \Theta\big(\log_2 n\big) \\
& = \Theta\big(\log_2n + 2^{\alpha(n)}\log_2\log_2n \big) + \Theta\big(\log_2n\big)
= \Theta\big(\log_2n\big)
\end{aligned}
\end{equation*}
%We bound the RHS by bounding $\runtime\left(\alpha_{\alpha(n)}(n)\right)$, and $\runtime\left(\alpha_i(n)\right)$ in general. Note that by $\alpha_2(n) = \left\lfloor \frac{n - 2}{2} \right\rfloor$, we have $\runtime(\alpha_2, n) \le 2\log_2 n$ for all $n$.
%Note that this bound can potentially be improved by further tightening the above inequalities. Although we do not obtain an exact asymptotic runtime similar 
%to Theorem~\ref{thm: inv-ack-hier-runtime-improved}, since this bound of $4^{\alpha(n)}\log_2n$ is strictly larger than the lower bound of $\log_2n$, it is still extremely small and can be bounded by simpler expressions such as $(\log_2n)^2$.
%Our result is an improved version of the inverse Ackermann function which runs in time $\bigO\big(4^{\alpha(n)}\log_2n\big)$.
%\begin{equation*}
%\alpha(n) \triangleq \begin{cases}
%0 & \text{if } n\le 1\\ 1 & \text{if } 2\le n\le 3 \\ 2 & \text{if } 4\le n\le 7 \\
%\W\alpha\left(\alpha_3,
%\alpha_3(n), 3, n\right) & \text{if } n\ge 8
%\end{cases} \quad \text{ where } \alpha_3 \triangleq \lambda m. \big(\lfloor \log_2(m+2) \rfloor - 2\big)
%\end{equation*}