Many functions on $\mathbb{R}$ are bijections and thus have an intuituve inverse.  
Functions on~$\mathbb{N}$ are often non-bijections and so their inverses
do not come as naturally.

\subsection{Upper Inverses, Expansions, and Repeatability}
%; indeed, simply defining an inverse can be a little subtle.
\begin{defn} \label{defn: inverse}
The \emph{upper inverse} of $F$, written $F^{-1}_{+}$, 
% {\color{red} $F^{-1}$, $F^{-1}_{\mathit{\shortuparrow}}$, $F^{-1}_{\upharpoonleft}$} 
is $\min\{m : F(m)\ge n\}$.\end{defn}
Notice that this is well-defined as long as $F$ is unbounded, \emph{i.e.} $\forall b.~\exists a.~ b \leq F(a)$.  However, as a notion of ``inverse,'' it really only makes sense if $F$ is strictly
increasing, \emph{i.e.} $\forall n,m.~n < m \Rightarrow F(n) < F(m)$, which is, in some sense, the analogue of injectivity in the discrete domain.

We call this function the ``upper inverse'' because, for strictly increasing functions like
addition, multiplication, and exponentiation, the upper inverse is the ceiling of the 
corresponding inverse functions on $\mathbb{R}$. We can characterize inverses more meaningfully as follows: 
\hide{
\color{red} It is reasonable to wonder about the floor.
\begin{defn} \label{defn: lower_inverse}
The converse \emph{lower inverse}, written $F^{-1}_{-}$,
is defined as $\max\{m : F(m)\le n\}$.
\end{defn}
Even if $F$ strictly increases, as $a[n]$ does for every $a\ge 2$, notice that the lower inverse will be undefined for $n < F(0)$, \emph{e.g.} $\{m : a[n]m \le 0 \} = \varnothing$ for $n\ge 3$.
Thus we focus on upper inverses (hereafter just ``inverses''), and discuss lower inverses in \cref{sec: discussion}. 
}%end hide 
\begin{thm} \label{thm: upp-inverse-rel}
	If $F:\mathbb{N}\to \mathbb{N}$ is increasing, then $f$ is the upper inverse of $F$ if and only if $\ \forall n, m.~ f(n)\le m \iff n \le F(m)$.
\end{thm}
\begin{proof}
Fix $n$. The sentence $\forall m.~ f(n)\le m \iff n\le F(m)$ implies: (1) $f(n)$ is a lower bound to $\{m: n \le F(m)\}$ and (2) $f(n)$ is itself in the set, since plugging in $m \triangleq f(n)$ yields $n\le F(f(n))$, which makes $f$ the upper inverse of $F$. Conversely, if $f$ is the upper inverse of $F$, we know $\forall m.~n\le F(m)\Rightarrow f(n)\le m$. Now, $\forall m \ge f(n)$.~$F(m)\ge F(f(n)) \ge n$ by increasing-ness, thus completing the proof.
\end{proof}
\begin{col}
If $F$ is strictly increasing, then $F^{-1}_{+} \circ F$ is
the identity function.
\end{col}
\begin{proof}
By ($\Leftarrow$) of Theorem~\ref{thm: upp-inverse-rel}, $F(n) \le F(n)$ implies 
$(F^{-1}_{+} \circ F)(n) \le n$.  By ($\Rightarrow$) of the same theorem, $(F^{-1}_{+} \circ F)(n) \le (F^{-1}_{+} \circ F)(n)$ implies $F(n) \le F \big((F^{-1}_{+} \circ F)(n)\big)$. $F$ is strictly increasing, so $n \le (F^{-1}_{+} \circ F)(n)$.
\end{proof}

Our setup for inverse requires increasing functions, and our definitions for 
hyperoperations/Ackermann use repeater.  Suppose $F$ is a strictly increasing function.
For a given $a$, is $\rf{F}{a}$ strictly increasing?  No!  For example, the identity function
$\li{id}$ is strictly increasing, but $\rf{\li{id}}{a}(n) = (\li{id} \circ \ldots \circ \li{id}) (a) = a$ is a constant function.  We need a little more.
\begin{defn}
A function $F:\mathbb{N}\to\mathbb{N}$ is an \emph{expansion} if $\forall n.~ F(n)\ge n$. Further, given $a\in \mathbb{N}$, an expansion $F$ is \emph{strict from} $a$ if \note{$\forall n \ge a.~ F(n)\ge n+1$}.
\end{defn}
If $a\ge 1$ and $F$ is an expansion \emph{strict from} $a$, we quickly get: 
$\forall n.~ \rf{F}{a}(n) = F^{(n)}(a) \ge a + n \ge 1 + n$. That is, $\rf{F}{a} \ $ is itself an expansion strict from $0$. 

\begin{defn} \label{rem: repeatable-subset}
$\forall a\ge 1$, if~$F$ is a strictly increasing function 
that is also a strict expansion from~$a$, then $\rf{F}{a}$ preserves 
repeatability.
We thus say that such a function $F$ is \emph{repeatable from} $a$, and denote the 
set of functions repeatable from $a$ as $\repeatable_a$. 
It is straightforward to see that $\forall s,t.~ s \le t \Rightarrow \repeatable_s \subseteq \repeatable_t $.
\end{defn}

\subsection{Contractions and the countdown operation}

Suppose $F \in \repeatable_a$ for some $a \ge 1$, and 
let $f \triangleq F^{-1}_{+}$, \emph{i.e.} the inverse of $F$.
Our goal is to use $f$ to compute an inverse to $\rf{F}{a}$.  
From the preceeding discussion we know that this inverse must exist, 
since $F \in \repeatable_a$ implies $\rf{F}{a} \in \repeatable_0$.  
For reasons that will be clear momentarily, we write this inverse as $\cdt{f}{a}$.  Now
fix $n$ and observe that $\forall m.~f^{(m)}(n)\le a \iff m \ge \cdt{f}{a}(n)$ since
\begin{equation} \label{eq: rf-upp-inv}
\begin{aligned}
\cdt{f}{a}(n)\le m & \iff n\le \rf{F}{a}(m) = F^{(m)}(a) \iff f(n)\le F^{(m-1)}(a) \\
& \iff f^{(2)}(n)\le F^{(m-2)}(a) \iff \ldots \iff f^{(m)}(n)\le a
\end{aligned}
\end{equation}
Moreover, setting $m = \cdt{f}{a}(n)$, we realize that 
$f^{(\cdt{f}{a}(n))}(n) \le a$.  
\textbf{Together these imply that $\cdt{f}{a}(n)$ is the minimum number of 
times~$f$ needs to be compositionally applied to $n$ before equalling or 
passing $a$.} 
In other words, count the length of the chain $\{n, f(n), f^{(2)}(n), \ldots\}$ that 
terminates as soon as we reach/pass $a$.  For this process to work, 
we need each chain link
to be strictly less than the previous, \emph{i.e.} $f$ must be a \emph{contraction}.
\begin{defn} \label{defn: contracting}
	A function $f : \mathbb{N} \to \mathbb{N}$ is a \emph{contraction} if $\forall n.~ f(n) \le n$. Given an $a \ge 1$, a contraction $f$ is 
	\emph{strict above} $a$ if $\forall n > a.~n > f(n)$. We denote the set of contractions by $\contract$ and the set of contractions strict above $a$ by $\contract_a$. Analogously to our observation in 
	Definition \ref{rem: repeatable-subset}, $\forall s\le t.~ \contract_s \subseteq \contract_t$.
\end{defn}
What kinds of functions have contractive inverses? Expansions, naturally:
\begin{thm} \label{thm: expansion-inv-contraction}
$\forall a\in \mathbb{N}.~F\in \repeatable_a \Rightarrow F^{-1}_+ \in \contract_a$.
\end{thm}
\begin{proof}
$\forall n.~F(n)\ge n \Rightarrow n \ge F^{-1}_+(n)$, so $F^{-1}_+ \in
\contract$. Note, $n > a \iff n-1\ge a$.
Since $F\in \repeatable_a$, $F(n-1)\ge n$ holds. 
Next, $n-1\ge F^{-1}_+(n) \Rightarrow n > F^{-1}_+(n)$. 
\end{proof}
This clarifies the inverse relationship between expansions strict from some $a$ and contractions strict above that same $a$. The inverse of an expansion's 
repeater exists and can be built \note{from its own inverse}, in a method formalized as \emph{countdown}.
\begin{defn} \label{defn: informal-countdown} \label{eq: countdown}
Let $f\in \contract_a$. The \textit{countdown to} $a$ of $f$, written 
$\cdt{f}{a}(n)$, is the smallest number of times $f$ needs to be applied to 
$n$ for the answer to equal or go below $a$. \emph{i.e.}, 
$\cdt{f}{a}(n) \triangleq \min\{m: f^{(m)}(n)\le a \}$.
\end{defn}
Inspired by Equation~\ref{eq: rf-upp-inv}, we provide a neat, algebraically manipulable logical sentence equivalent to Definition~\ref{eq: countdown}, which is more useful later in our paper:
\begin{col} \label{col: cdt-repeat}
If $a \in \mathbb{N}$ and $f\in \contract_{a}$, then $\forall n, m.~ \cdt{f}{a}(n)\le m \iff f^{(m)}(n)\le a$.
\end{col}
\begin{proof}
	Fix $a$ and $n$. The interesting direction is $(\Rightarrow)$. Suppose $\cdt{f}{a}(n)\le m$, we get $f^{(m)}(n)\le f^{(\cdt{f}{a}(n))}(n)$ due to $f\in \contract$, and $f^{(\cdt{f}{a}(n))}(n)\le a$ due to Definition~\ref{eq: countdown}.
\end{proof}
%\begin{thm} \label{thm: upp-inv-cdt-rf}
%	For all $a\in \mathbb{N}$, if $f\in \contract_{a}$ is the upper inverse of $F: \mathbb{N}\to \mathbb{N}$, then $\cdt{f}{a}$ is the upper inverse of $\rf{F}{a}$.
%\end{thm}
Another useful result is the recursive formula for \emph{countdown}:
\begin{thm} \label{thm: cdt-recursion}
	For all $a\in \mathbb{N}$ and $f\in \contract_{a}$, $\cdt{f}{a}$ satisfies:
	\begin{equation*}
	\cdt{f}{a}(n) = \begin{cases}
	0 & \text{ if } n \le a \\ 1 + \cdt{f}{a}(f(n)) & \text{ if } n \ge a + 1
	\end{cases}
	\end{equation*}
\end{thm}
\begin{proof}
When $n \le a$, use Corollary~\ref{col: cdt-repeat} as follows: 
$n = f^{(0)}(n)\le a \iff \cdt{f}{a}(n)\le 0$. 
When $n\ge a+1$, proceed by antisymmetry. 
Define $m \triangleq \cdt{f}{a}(f(n))$, and note that 
Corollary~\ref{col: cdt-repeat} gives 
$\cdt{f}{a}(n)\le 1+m \iff f^{(1+m)}(n) \le a$.
Next, a simple expansion of the last clause gives $f^{(m)}(f(n)) \le a$,
and unfolding the definition of~$m$ shows that this last clause is true.
Now since $n\ge a+1$, we have $\cdt{f}{a}(n)\ge 1$ by the above. 
Define $p \triangleq \cdt{f}{a}(n) - 1$. 
It remains to show $\cdt{f}{a}(f(n))\le p \Rightarrow f^{(p)}(f(n))\le a \Rightarrow f^{(p+1)}(n)\le a$, which holds by unfolding $p$'s definition.
\end{proof}

%\begin{thm} \label{thm: cdt-contr-0}
%	For all $a\ge 1$, if $f\in \contract_{a}$, then $\cdt{f}{a}\ \in \contract_{c} \ \forall c$.
%\end{thm}
%\begin{proof}
%	Firstly we show that $\cdt{f}{a}$ is a contraction, namely showing $\cdt{f}{a}(n)\le n \ \forall n$, which has already been proved in the proof of \cref{thm: cdt-repeat}. To show $\cdt{f}{a}$ is strict from $1+c$, it suffices to show it is strict from $1$, equivalently $\cdt{f}{a}(n) \le n - 1 \ \forall n\ge 1$. By \cref{thm: cdt-repeat}, we need to show $f^{(n-1)}(n)\le a$. Assume the converse, then:
%	$$ n \ge 1 + f(n)\ge 2 + f(f(n)) \ge \cdots \ge (n-1) + f^{(n-1)}(n) $$
%	Thus $f^{(n-1)}(n)\le 1 \le a$, a contradiction. The theorem then follows.
%\end{proof}

\subsection{A Structurally Recursive Computation for Countdown}

The higher-order repeater function is well-defined for all input functions, 
even those not in $\repeatable_a$ (although for such functions it may not 
be useful), and so is easy to define
in Coq as shown in \S\ref{sec: countdown-repeater}. In contrast, a 
\emph{countdown} only exists for certain functions, most conveniently 
contractions, which makes it a little harder to encode into Coq. 
Our strategy is to define a \emph{countdown worker} which is written with only
structural recursion and is thus more palatable to Coq, and then 
prove that this worker computes the countdown when passed a contraction.

\begin{defn} \label{defn: countdown-worker}
For any $a\in \mathbb{N}$ and $f: \mathbb{N}\to \mathbb{N}$, the \emph{countdown worker} to $a$ of $f$ is a function $\cdw{f}{a}\ : \mathbb{N}^2\to \mathbb{N}$ such that:
\begin{equation*}
\cdw{f}{a}(n, b) = \begin{cases}
0 & \text{if } b = 0 \vee n\le a \\ 1 +\cdw{f}{a}(f(n), b-1) & \text{if } b \ge 1 \wedge n > a
\end{cases}
\end{equation*}
\end{defn}
The \emph{countdown worker} operates on two arguments: 
the \emph{true argument} $n$, for which we want to simulate 
\emph{countdown to} $a$, 
and the \emph{budget} $b$, the maximum number of times we shall attempt 
to compositionally apply $f$ on the input before giving up. 
If the input goes below or equals $a$ after $k$ applications, \emph{i.e.} $f^{(k)}(n) \le a$, we return the count $k$. If the budget is exhausted (\emph{i.e.} $b = 0$) while the result is still above $a$, we fail by returning the original budget. This definition is admittedly inelegant, but it can clearly be written as a Coq \li{Fixpoint}:
\begin{lstlisting}
`\href{https://github.com/inv-ack/inv-ack/blob/6099297c6ab0e16d14b037fb5ed600c4d22818f6/countdown.v#L94-L100}{Fixpoint countdown\_worker}` (a : nat) (f : nat -> nat) (n k : nat) : nat :=
match k with
 | 0    => 0
 | S k' => match (n - a) with
           | 0 => 0
           | _ => S (countdown_worker a f (f n) k') 
           end
 end.
\end{lstlisting}
Given an $f$ that is a contraction strict from $a$, 
and given a sufficient budget, \emph{countdown worker} 
will compute the correct \emph{countdown} value.  
Careful readers may have noticed that \emph{budget} is similar to 
\emph{gas}, which we discussed in ~\S\ref{sec:incfuncinv} 
and dismissed for potentially 
being too computationally expensive to calculate. 
Budget is actually a refinement of gas because 
we support its use with a method to calculate it in constant time.
In fact, we will soon show that a budget of $n$ is sufficient. 
This lets us define \emph{countdown} in Coq for the first time:
\begin{defn} \label{defn: countdown}
Redefine $\cdt{f}{a}(n) \triangleq \cdw{f}{a}(n, n)$.
\begin{lstlisting}
`\href{https://github.com/inv-ack/inv-ack/blob/6099297c6ab0e16d14b037fb5ed600c4d22818f6/countdown.v#L104}{Definition countdown\_to}` a f n := countdown_worker a f n n.
\end{lstlisting}
\end{defn}
%Before beginning, let us clarify that the definition of $\mathbb{N}$ and operations on $\mathbb{N}$ in Gallina follow the Presburger Arithmetic \cite{presburger}, which despite being weaker than Peano Arithmetic, is a decidable theory. The Coq standard library includes \texttt{Omega}, an extensive listing of provable facts about $\mathbb{N}$ in Presburger Arithmetic, including everything used in this paper, most notably the law of excluded middle for comparisons on $\mathbb{N}$:
%\begin{equation*}
%(n \le m) \vee (m + 1 \le n) \ \ \forall n, m
%\end{equation*}
%, which is provable without the actual law of excluded middle in classical logic. This enables us to prove all results in this paper with Coq's baseline intuitionistic logic. Readers can refer to the \Cref{appendix} for Coq versions of the proofs.
%in which the most operations agree with usual operations on $\mathbb{Z}$, except subtraction, which is defined as:
%\begin{lstlisting}
%Fixpoint sub (n m : nat) : nat :=
%match n with
%| 0 => n
%| S k => match m with
%| 0 => n
%| S l => sub k l end
%end.
%\end{lstlisting}
%Essentially $\li{sub} \ n \ m = \max\{n - m, 0\}$. We will use this subtraction for the rest of the paper.

%Firstly, we begin with a lemma asserting the existence of the countdown value itself. Although its existence is guaranteed by the well-ordering principle of $\mathbb{N}$, we will achieve better by proving it in intuitionistic logic.
%
%\begin{lem} \label{lem: contract-repeat-threshold}
%	For all $a, n\in\mathbb{N}$ and $f\in \contract_{a}$,
%	\begin{equation}
%	\exists m : \left(f^{(m)}(n) \le a \right) \wedge \left(f^{(l)}(n)\le a \Rightarrow m \le l \ \forall l \right)
%	\end{equation}
%\end{lem}
%\begin{proof}
%  Fix $n$ and observe that if $n\le a$, $m = 0$ is the desired choice since $ f^{(0)}(n) = n \le a \ \text{ and } \ 0 \le l \ \forall l $.
%	Consider only when $a\le n$, we can define $c$ such that $n = a + c$. We prove the following statement by induction on c:
%	\begin{equation*}
%	P(c) \triangleq \exists m : \left(f^{(m)}(n) \le n - c \right) \wedge \left(f^{(l)}(n)\le n - c \Rightarrow m \le l \ \forall l \right)
%	\end{equation*}
%	under assumptions $f\in \contract_{n-c}$ and $c\le n$.
%	\begin{enumerate}[leftmargin=*]
%		\item \textit{Base case.} The case $c = 0$ implies $n = a$, which has been proven above.
%		\item \textit{Inductive step.} Suppose $P(c)$ is proved with witness $m_c$. Note that the assumptions are now $f\in \contract_{n-c}$ and $c+1\le n$, there are two cases:
%		\begin{itemize}[leftmargin=*, label={--}]
%			\item $f^{\left(m_c\right)}(n) = n - c$. Then $f^{\left(m_c+1\right)}(n) \le n - c - 1$. Let $m_{c+1} = m_c + 1$, for all $l$:
%			\begin{equation*}
%			f^{(l)}(n)\le n - c - 1 < f^{\left(m_c\right)}(n) \Rightarrow l > m_c \Rightarrow l \ge m_{c+1}
%			\end{equation*}
%			\item $f^{\left(m_c\right)}(n) \le n - c - 1$. Let $m_{c+1} = m_c$, for all $l$:
%			\begin{equation*}
%			f^{(l)}(n)\le n - c - 1\le n - c \overset{P(c)}{\Rightarrow} l\ge m_{c} = m_{c+1}
%			\end{equation*}
%		\end{itemize}
%		In any cases, we can find a witness $m_{c+1}$ for $P(c+1)$. Thus the proof is complete by induction.\vspace*{-\baselineskip}
%	\end{enumerate}
%\end{proof}

\begin{lem} \label{lem: cdt-init}
	$\forall n\le a.~\cdw{f}{a} (n, b) = 0$.
\end{lem}
\begin{proof}
This simple fact follows directly from Definition \ref{defn: countdown-worker}.
\end{proof}
Next, we show the internal working of \emph{countdown worker} at the $\text{i}^\text{th}$ recursive step, including the accumulated result $1+i$, the current input $f^{(1+i)}(n)$, and the current budget $b-i-1$.
\begin{lem} \label{lem: cdt-intermediate}
	$\forall a, n, b, i \in \mathbb{N}.~\forall f \in \contract$. such that $i < b$ and $a < f^{(i)}(n)$:~
	\begin{equation}  \label{eq: cdt-intermediate}
	\cdw{f}{a}(n, b) = 1 + i + \cdw{f}{a}\left(f^{1+i}(n), b - i - 1\right)
	\end{equation}
\end{lem}
\begin{proof}
	Fix $a$. We proceed by induction on $i$. Then define $P(i)$ as
	\begin{equation*}
	P(i) \triangleq \forall b, n.~ b\ge i+1 \Rightarrow f^{(i)}(n) > a \Rightarrow \cdw{f}{a}(n, b) = 1 + i + \cdw{f}{a}\left(f^{1+i}(n), b - i - 1\right)
	\end{equation*}
	\begin{itemize}[leftmargin=*]
		\item \textit{Base case.} For $i = 0$, our goal $P(0)$ is:
		$\cdw{f}{a}(n, b) = 1 + \cdw{f}{a}\left(f(n), b - 1\right)$
		where $b \ge 1$ and $f(n)\ge a+1$. This is straightforward.
		\item \textit{Inductive step.} Assume $P(i)$ has been proved. Then $P(i+1)$ is
		\begin{equation*}
		\cdw{f}{a}(n, b) = 2 + i + \cdw{f}{a}\left( f^{2+i}(n), b - i - 2 \right)
		\end{equation*}
		where $b \ge i+2$ and $f^{1+i}(n) \ge a+1$. This also implies $b\ge i+1$ and $\displaystyle f^{(i)}(n) \ge f^{1+i}(n)\ge a+1$ by $f\in \contract$, and so $P(i)$ holds. It suffices to prove:
		\begin{equation*}
		\cdw{f}{a}\left(f^{1+i}(n), b-i-1\right) = 1 + \cdw{f}{a}\left( f^{2+i}(n), b-i-2 \right)
		\end{equation*}
		This is in fact $P(0)$ with $(b, n)$ substituted for $\left(b-i-1, f^{(1+i)}(n)\right)$. Since $f^{(1+i)}(n) \ge a+1$ and $b-i-1\ge 1$, the above holds and $P(i+1)$ follows.
	\end{itemize}
\end{proof}
Now it is time to prove the correctness of \emph{countdown}, \emph{i.e.} that our computational Definition~\ref{defn: countdown} computes the value originally specified in Definition~\ref{defn: informal-countdown}.
\begin{thm} \label{thm: cdt-repeat}
	$\forall a \in \mathbb{N}.~ \forall f\in \contract_{a}$, we have 
%	\begin{equation} \label{eq: cdt-minimum}
$	\forall n.~ \cdt{f}{a}(n) = \min\left\{ i : f^{(i)}(n) \le a \right\} $.
%	\end{equation}
%	Or equivalently,
%	\begin{equation} \label{eq: cdt-repeat}
%	\cdt{f}{a}(n) \le k \iff f^{(k)}(n) \le a \ \ \forall n, k
%	\end{equation}
\end{thm}
\begin{proof}
%	First, to see why \eqref{eq: cdt-minimum} and \eqref{eq: cdt-repeat} are equivalent, we rewrite \eqref{eq: cdt-minimum} in the following way:
%	$$ \left(f^{(\cdt{f}{a}(n))}(n) \le a\right) \wedge \left(f^{(l)}(n) \le a \Rightarrow \cdt{f}{a}(n) \le l \ \ \forall l\right) \ \ \forall n$$
%	To prove $\eqref{eq: cdt-minimum} \Rightarrow \eqref{eq: cdt-repeat}$, it suffices to show
%	$$ \cdt{f}{a}(n) \le l \Rightarrow f^{(l)}(n) \le a \ \ \forall l $$
%	, which holds due to the fact $\displaystyle f^{(l)}(n) \le f^{(\cdt{f}{a}(n))}(n) \le a$ by $f\in \contract$. To prove $\eqref{eq: cdt-repeat}\Rightarrow \eqref{eq: cdt-minimum}$, it suffices to show $\displaystyle f^{(\cdt{f}{a}(n))}(n) \le a$
%	, which in turn holds by substituting $k$ by $\cdt{f}{a}(n)$ in \eqref{eq: cdt-repeat}. Thus $\eqref{eq: cdt-minimum}\iff \eqref{eq: cdt-repeat}$ and we need only to prove \eqref{eq: cdt-minimum}.
Since $f\in \contract_{a}$ and $\mathbb{N}$ is well-ordered, 
let $m = \min\big\{i : f^{(i)}(n)\le a\big\}$.\footnote{We prove the existence 
of the min in Coq’s intuitionistic logic \href{https://github.com/inv-ack/inv-ack/blob/6099297c6ab0e16d14b037fb5ed600c4d22818f6/countdown.v\#L125-L150}{here} in our codebase.} Our goal becomes 
$\cdt{f}{a}(n) = m$. Note that this setup gives us: 
	\begin{equation}
	\left(f^{(m)}(n) \le a\right) \label{eq: cdt-repeat-tmp} \wedge
	 \left(\forall k.~f^{(k)}(n)\le a \Rightarrow m \le k\right)
	\end{equation}
If $m = 0$, then $n = f^{(0)}(n)\le a$. So $\cdt{f}{a}(n) = \cdw{f}{a}(n, n) = 0 = m$ by Lemma~\ref{lem: cdt-init}.
When $m > 0$, our plan is to apply Lemma~\ref{lem: cdt-intermediate} to get
	\begin{equation*}
	\cdt{f}{a}(n) = \cdw{f}{a}(n, n) = m + \cdw{f}{a}\left(f^{(m)}(n), n-m\right),
	\end{equation*}
and then use Lemma~\ref{lem: cdt-init} over (\ref{eq: cdt-repeat-tmp})'s first conjunct to conclude $\cdt{f}{a}(n) = m$. It suffices to prove the premises of Lemma~\ref{lem: cdt-intermediate}: $a < f^{(m-1)}(n)$ and $m-1 < n$.
	
	The former follows by contradiction: if $f^{(m-1)}(n) \le a$, (\ref{eq: cdt-repeat-tmp})'s second conjunct implies $m\le m-1$, which is impossible for $m > 0$. The latter then easily follows by $f\in \contract_{a}$, since
	$n \ge 1 + f(n) \ge 2 + f(f(n)) \ge \cdots \ge m + f^{(m)}(n)$.
	\linebreak Therefore, $\cdt{f}{a}(n) = m$ in all cases, which completes the proof.
\end{proof}
Theorem~\ref{thm: cdt-repeat} and (\ref{eq: rf-upp-inv}) establish the correctness of the Coq definitions of \emph{countdown worker} and \emph{countdown}, thereby justifying our budget of $n$ and our unification of 
Definitions \ref{defn: informal-countdown} and \ref{defn: countdown}. We wrap everything together with the following theorem:
\begin{thm} \label{thm: cdt-inv-rf}
	$\forall F\in \repeatable_a.~f\triangleq F^{-1}_+$ satisfies $f \in \contract_{a}$ and $\displaystyle \cdt{f}{a} \ = \left(\rf{F}{a}\ \right)^{-1}_+$. Furthermore, if $a\ge 1$, then $\rf{F}{a}\ \in \repeatable_0$ and $\cdt{f}{a}\ \in \contract_0$.
\end{thm}
\begin{proof}
	By Theorem~\ref{thm: expansion-inv-contraction}, $f\triangleq F^{-1}_+ \in \contract_a$. 
	\eqref{eq: rf-upp-inv} and Corollary~\ref{col: cdt-repeat}
	then show $\cdt{f}{a}\ = \left(\rf{F}{a}\ \right)^{-1}_+$.
	Now if $a\ge 1$, a simple induction shows that $F^{(n)}(a)\ge a + n\ge 1 + n$, so $\rf{F}{a}\ \in \repeatable_0$, hence $\cdt{f}{a} \ = \left(\rf{F}{a}\ \right)^{-1}_+ \in \contract_0$ by Theorem~\ref{thm: expansion-inv-contraction}.
\end{proof}

