\newcommand{\ackt}{\ensuremath{\hat{\alpha}}}

\subsection{The Value of a Linear-Time Solution to the Hierarchy}

Our functions' linear runtimes can be understood in two distinct but
complementary ways.  A runtime less than the bitlength is impossible
without prior knowledge of the size of the input.  Accordingly, in
an information-theory or pure-mathematical sense, our definitions are
optimal up to constant factors.  And of course in practice, linear-time
solutions are highly usable in real computations.

Sublinear solutions are possible with \emph{a priori} knowledge about
the function and bounds on the inputs one will receive.
An extreme case is $\alpha(n)$, which has value $4$ for all practical
inputs greater than $61$. Accordingly,
this function can be inverted in $O(1)$ in practice.  That said, 
such solutions require external knowledge of the problems and
lookup tables within the algorithm to store precomputed
values, and thus fall more into the realm of engineering than mathematics. 

\begin{rem}
Rather surprisingly, for binary-encoded numbers, it appears we can invert the Ackermann function asymptotically faster than we can calculate the base-3 logarithm. 
\end{rem}

\subsection{Benchmarking our Inverses}

\begin{table}[t]
	\begin{centermath}
		\begin{array}{cl@{\hspace{8em}}c}
			& \li{n} & \text{time in }\li{OCaml} \\
			\cline{2-3}
			\cline{2-3}
			\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\li{n:nat}}}} 
			& 100 & 0.000020 \\
			& 1000 & 0.000926 \\
			& 10000 & 0.001450 \\
			& 100000 & 0.016406 \\	
			\cline{2-3}
			\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{\li{n:N}}}} 
			& 2^{100} & 0.000012 \\
			& 2^{1000} & 0.000082 \\
			& 2^{10000} & 0.000959 \\
			& 2^{100000} & 0.014332 \\
		\end{array}
	\end{centermath}
	\caption{Benchmarking our results \\ (all times in seconds).}
	\label{table:benchmarking}
	\vspace{0em}
\end{table}

As shown in Table~\ref{table:benchmarking}, our functions are very rapid in practice.
Extracting our code to OCaml is straightforward: we import Coq's built-in \li{Extraction} module and then execute \li{Recursive Extraction inv_ack_linear.} We then benchmark our results using a machine running i5 2.4GHz with 8GB RAM.  The displayed times are
the average of three runs, and we suspect that garbage collection is the largest cause 
of variations in the wall-clock times.  Our functions run quickly
in Gallina as well, but exhibit unexpectedly rapid times when the inputs are 
large because Coq optimizes large \li{nat}s under the hood.

%we show the results of 
%extracting our linear-time inverse Ackermann functions to OCaml using Coq's built-in 
%\li{Extraction} utility and then running a few benchmarks.
%We made no further optimizations in OCaml, and the timings above
%are $3$-time averages of wallclock timings on 

\subsection{The Two-Parameter Inverse Ackermann Function}
Some authors~\cite{chazelle,tarjan} prefer a two-parameter inverse Ackermann function.
\begin{defn} \label{defn: 2para-alpha}
	The two-parameter inverse Ackermann function is defined as:
	\begin{equation} \label{eq: tmp-2para-alpha}
	\ackt (m, n) \triangleq \min\left\{i \ge 1 : \Ack\left(i, \left\lfloor \frac{m}{n} \right\rfloor \right)\ge \log_2n \right\}
	\end{equation}
\end{defn}
\noindent Note that $\ackt(n, n) \neq \alpha(n)$. However,
it is straightforward to modify our techniques to compute $\ackt(m, n)$.
\hide{This function arises from deep runtime analysis of the disjoint-set data structure. Tarjan \cite{tarjan} showed that, in the disjoint-set data structure, the time required $t(m,n)$ for a sequence of $m$ \textsc{\color{magenta}FIND}s intermixed with $n-1$ \textsc{\color{magenta}UNION}s (such that $m \geq n$) is bounded as: $k_{1}m\cdot\alpha(m,n) \leq t(m,n) \leq k_{2}m\cdot\alpha(m,n)$. In graph theory, Chazelle \cite{chazelle} showed that the minimum spanning tree of a connected graph with $n$ vertices and $m$ edges can be found in time $O(m\cdot\alpha(m,n))$. Computing this function is in fact easier than $\alpha(n)$, as when $m$ and $n$ are given, we are reduced to finding the minimum $i\ge 1$ such that $\Ack_i(s)\ge t$ for $s, t$ fixed, which can be done with the following variant of the \emph{inverse Ackermann worker}.
}
\begin{defn} \label{defn: 2para-inv-ack-worker}
	The {two-parameter inverse Ackermann worker},
	written $\ackt^{\W}$, is a function $\mathbb{N}^4\to \mathbb{N}$, defined by:
	\hide{$(\mathbb{N}\to \mathbb{N}) \times \mathbb{N}^3\to \mathbb{N}$ such that for all $n, k, b\in \mathbb{N}$ and $f:\mathbb{N}\to \mathbb{N}$:}
%	\begin{equation} \label{eq: 2para-inv-ack-worker-recursion}
%	\ackt^{\W}(f, n, k, b) = \begin{cases}
%	0 & \text{if } b = 0 \vee n\le k \\ 1 + \ackt^{\W}\big(\cdt{f}{1}\circ f , \cdt{f}{1}(n), k, b-1\big) & \text{if } b \ge 1 \wedge n \ge k+1
%	\end{cases}
%	\end{equation}
  \begin{equation} \label{eq: 2para-inv-ack-worker-recursion}
  \begin{aligned}
  & \ackt^{\W}(f, n, k, b) \\
  & \triangleq \begin{cases}
  0 & \text{if } b = 0 \vee n\le k \\ 1 + \ackt^{\W}\big(\cdt{f}{1}\circ f , \cdt{f}{1}(n), k, b-1\big) & \text{if } b \ge 1 \wedge n \ge k+1
  \end{cases}
  \end{aligned}
  \end{equation}
\end{defn}
%Similar to the one-parameter version, the following theorem establishes the correct setting for $\W\alpha_2$ to compute $\alpha(m, n)$.
%\begin{thm}
%	$\displaystyle \ackt(m, n) = 1 + \ackt^{\W}\left(\alpha_1, \alpha_1\big(\lceil\log_2n \rceil\big), \left\lfloor \frac{m}{n} \right\rfloor, \lceil\log_2n \rceil \right)$.
%\end{thm}
% Edited by Linh
\begin{thm} We can calculate $\ackt(m, n)$ from $\ackt^{\W}$:
	\begin{equation*}
	\displaystyle \forall m,n.~\ackt(m, n) = 1 + \ackt^{\W}\left(\alpha_1, \alpha_1\big(\lceil\log_2n \rceil\big), \left\lfloor \frac{m}{n} \right\rfloor, \lceil\log_2n \rceil \right).
	\end{equation*}
\end{thm}
\noindent We mechanize the above for both \href{https://github.com/inv-ack/inv-ack/blob/7270e64a2600b771f2b1b1b151f7d13fb2ae6c97/inv_ack.v#L245-L248}{\color{blue}unary} and \href{https://github.com/inv-ack/inv-ack/blob/7270e64a2600b771f2b1b1b151f7d13fb2ae6c97/bin_inv_ack.v#L222-L228}{\color{blue}binary} inputs.

\hide{
	\begin{proof}[Proof Sketch]
		It is easy to prove in a similar fashion to \cref{lem: inv-ack-worker-intermediate} that for all $n, b, k$ and $i$, if $\alpha_i(n) > k$ and $b > i$, then
		\begin{equation*}
		\W\alpha_2\big(\alpha_1, \alpha_1(b), k, b\big) = i + \W\alpha_2\big(\alpha_{i+1}, \alpha_{i+1}(n), k, b - i\big)
		\end{equation*}
		Now let $k \triangleq \lfloor m/n \rfloor$, $b \triangleq \lceil \log_2n \rceil$ and $l \triangleq \min\big\{i : \alpha_i(b)\le k\big\}$, which exists because $\Ack(i, \cdot)$ increases strictly with $i$. Then $\alpha(m, n) = \max{1, l}$. If $l = 0$, $\alpha_1(b) \le \alpha_0(b) \le k$, so $\W\alpha_2\big(\alpha_1, \alpha_1(b), k, b\big) = 0$, as desired. If $l \ge 1$,
		\begin{equation*}
		1 + \W\alpha_2\big(\alpha_1, \alpha_1(b), k, b\big)
		= 1 + l - 1 + \W\alpha_2\big(\alpha_l, \alpha_l(b), k, b-l\big) = l
		\end{equation*}
		Here $b\ge l$ due to the fact that $\Ack(b, k)\ge b$, so $\alpha_b(b)\le k$. This completes the proof.
\end{proof}}%end hide

\subsection{Implementing our Functions in Isabelle/HOL}

To show that our functions are independent of any 
Coq idiosyncrasies,
we implemented our functions in Isabelle/HOL.  Despite having 
no previous experience with Isabelle/HOL, it took us
only 2 hours to define our inverse Ackermann function.
Figure~\ref{fig:standalone_isabelle} is an Isabelle translation of 
Figure~\ref{fig:standalone}.

\lstset{style=isaStyle}
\begin{figure}
\begin{lstlisting}
theory inv_ack_standalone
  imports "HOL-Library.Log_Nat" HOL.Divides
begin

primrec cdn_wkr :: "(nat => nat) => nat => 
                   nat => nat => nat" where
  "cdn_wkr f a n 0 = 0" |
  "cdn_wkr f a n (Suc k) = 
    (if n <= a then 0 
  	           else Suc (cdn_wkr f a (f n) k))"

fun countdown_to :: "(nat => nat) => nat => 
                    nat => nat" where
  "countdown_to f a n = cdn_wkr f a n n"

primrec inv_ack_wkr :: "(nat => nat) => nat => 
                       nat => nat => nat" where
  "inv_ack_wkr f n k 0 = k" |
  "inv_ack_wkr f n k (Suc b) = 
  	(if n <= k then k
  	else let g = (countdown_to f 1) in
    inv_ack_wkr (g o f) (g n) (Suc k) b)"

fun inv_ack_linear :: "nat => nat" where
  "inv_ack_linear 0 = 0" |
  "inv_ack_linear (Suc 0) = 0" |
  "inv_ack_linear (Suc (Suc n)) = 
  	inv_ack_wkr (`$\lambda$` x. (x - 2)) n 1 (Suc n)"
 
end
\end{lstlisting}
\caption{A linear-time Isabelle computation of the inverse Ackermann function (inputs  in unary, \emph{i.e.} \li{nat}).}
\label{fig:standalone_isabelle}
\end{figure}

%\noindent Although we have not done the proofs about our functions in Isabelle, they only require first-order mathematics and so should pose no difficulties.

\begin{figure}
\begin{lstlisting}
definition ln :: "real => real" where
  "ln x = (THE u. exp u = x)"

definition log :: "[real,real] => real" where
  "log a x = ln x / ln a"

definition floorlog :: "nat => nat => nat" where
  "floorlog b a = (if a > 0 `$\wedge$` b > 1 
                   then nat `$\lfloor$`log b a`$\rfloor$` + 1 else 0)"

lemma compute_floorlog[code]:
  "floorlog b x = (if x > 0 `$\wedge$` b > 1 then 
              floorlog b (x div b) + 1 else 0)"
 by (* proof elided *)
\end{lstlisting}
\caption{The standard Isabelle technique for extractable discrete logarithm.}
\label{fig:isabelle_hack}
\end{figure}

Furthermore, our techniques are as \emph{applicable} to Isabelle as they are to Coq:
Isabelle offers \li{floorlog}, a discrete logarithm with arbitrary base, but 
\li{floorlog} is defined using~\li{ln}, the continous logarithm on $\mathbb{R}$.
This does not yield a computable function, and so the relevant 
Isabelle standard library
uses a ``hack'' (in the form of a \li{[code]} annotation) to generate a computable equivalent for extraction.
In Figure~\ref{fig:isabelle_hack} we present relevant extracts from the Isabelle standard libraries~\cite{isastan2013, isastan2019}.

This hack requires a pre-developed computational strategy 
for the \li{[code]} extraction-substitution lemma: there ain't no such thing 
as a free lunch~\cite{moonmistress}. Thus, \li{[code]}-extracting the inverse 
Ackermann would require the Isabelle/HOL functions presented in Figure~\ref{fig:standalone_isabelle} or some equivalent thereof. 
Further, we found no definitions of $\log^*$ 
or the inverse Ackermann function in the Isabelle documentation. Our technique 
gives directly extractable solutions to the entire hierarchy.
\lstset{style=myStyle}

% https://www.isa-afp.org/browser_info/current/AFP/IEEE_Floating_Point/Log_Nat.html
% https://isabelle.in.tum.de/website-Isabelle2013/dist/library/HOL/Log.html
